Reliable feature point matching

### 2013-08-27 text/main/ch3_matching_methods.tex
Added ball tree to github

In this chapter we will describe the algorithms used in the experiements
in detail. We start by introducing the proposed algorithms \emph{Mirror 
Match (MM)} and \emph{Mirror Match with Clustering (MMC)} and continue 
with the reference algorithms used later for comparison, \emph{Ratio}, 
\emph{Spectral} and \emph{Isomatch}.

\section{Proposed Matching Methods}
\label{S:MatchingMethods}
In this section we describe the two proposed algorithms for matching 
feature points \emph{Mirror Match
(MM)} and \emph{Mirror Match with Clustering (MMC)}.

@@

\begin{figure}[h]
    \centering%
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_pitts_source}
            \caption{Source image pair}
            \label{fig:pitts_source}
        \end{subfigure}%
        ~ %
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_pitts_keypoints}
            \caption{Feature points}
            \label{fig:pitts_keypoints}
        \end{subfigure}%
        \\ %add desired spacing between images, e. g. ~, \quad, 
        %\qquad (or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match_off}
            \caption{\emph{Ratio}}
            \label{fig:unique}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, 
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match_with_pruned}
            \caption{\emph{MM} intermediate result}
            \label{fig:within}
        \end{subfigure}%
        \\ %add desired spacing between images, e. g. ~, \quad, 
        %\qquad (or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match}
            \caption{\emph{MM} final result}
            \label{fig:without}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, 
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_partition}
            \caption{\emph{MMC} Partition Example}
            \label{fig:pitts_partition}
        \end{subfigure}%
    \caption{Feature matching with \emph{MM} and \emph{MMC}. Dots 
        represent feature points; green/red lines indicate 
        correct/incorrect matches, respectively; black lines represent 
        edges in the feature graph.  (c) Result of \emph{Ratio} 
        matching.  (d) All matches found by \emph{MM}, including 
        intra-image matches.  (e) Final \emph{MM} result.  (f) Example 
    of a graph partition of feature points after clustering in 
\emph{MMC}, which includes similar feature points from different sides 
of a building.}%
    \label{fig:comparemirror}%
\end{figure}%

The central idea behind \emph{MM} is to match features of $n$ images by 
taking every feature from all $n$ images and matching them against every 
other feature from the same set. We then discard the correspondences 
that match two points within the same image. Algorithm~\ref{alg-mm} 
details the implementation of \emph{MM}.

\begin{algorithm}[htb]
\caption{Mirror Match (\emph{MM})}
\label{alg-mm}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Acquisition Stage
	\State $features\gets features \cup getFeatures(I_i)$
\EndFor
\ForAll{$f_i \in features$} \Comment Matching Stage
	\State $f_m,f_n \gets get2NNs(f_i, features \setminus 
	\left\{f_i\right\})$
    \State $ratio \gets d(f_i, f_m) / d(f_i, f_n)$
	\If{$ratio < t$}
		\State $matches_{init} \gets matches_{init} \cup \left(f_i, f_m\right)$
	\EndIf
\EndFor
\ForAll{$\left(f_i, f_j \right) \in matches_{init}$} \Comment Filter 
Stage
\If{$\left(f_j, f_i \right) \in matches_{init} \wedge getImg(f_i) \neq 
getImg(f_j) \wedge \left(f_j, f_i\right) \not\in matches_{final}$}
		\State $matches_{final} \gets (f_i, f_j)$
	\EndIf
\EndFor \\
\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

In the \emph{acquisition stage} we extract local feature points and 
obtain their descriptors from the set of images. The algorithm is not 
designed to make use of any feature extractor in particular, but for the
experiments (Chapter~\ref{C:Experiments}) the \emph{SIFT} feature 
extractor and descriptor has been used. In the \emph{matching} stage 
these features are matched using $k$-nearest neighbors.  The distance 
function $d(f_i, f_j)$ between features is calculated according to the 
feature descriptor. For \emph{SIFT} we use the cosine similarity between 
the two descriptor vectors, while other descriptors might use other 
measures\footnote{\emph{BRIEF} for example uses the hamming measure to 
speed up the computation.}.  For any given feature $f_i$ the two closest 
neighbors are returned, and we calculate the ratio between them as 
proposed in \cite{lowe2004sift}\footnote{In Part~\ref{ss:ratio} this 
calculation and other details of \emph{Ratio} will be discussed.}.  Any 
correspondence with a ratio above the threshold supplied will be 
discarded. Finally in the filter stage we check that matches are from 
different images and discard all matches that are not symmetric. That 
is, if a feature $f_i$, is best matched to another feature $f_j$, we 
would want for $f_j$ to be best matched to $f_i$ too.
This isn't always the case, for example if $f_i$ does not have a true 
correspondance in the other image, its closest neighbor $f_j$ might 
likely be better matched with a different feature point.

Figure~\ref{fig:comparemirror} illustrates the benefits of \emph{MM} 
using an example image pair from the \emph{Gallagher} dataset 
\cite{gallagher2008}.
With \emph{Ratio} (Figure~\ref{fig:unique}), many incorrect matches occur 
in the fence towards the bottom of the image.
When we match all feature points together, many of these incorrect 
matches are eliminated, because points in the fence match with other 
points in the fence in the same image (Figures~\ref{fig:within} and
\ref{fig:without}).

@@

In contrast to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering feature points by 
similarity. This process yields partitions of similar feature points 
that we can match using the same approach as \emph{MM}.  
Algorithm~\ref{alg-mmc} shows the implementation of \emph{MMC}.

\begin{algorithm}[htb]
\caption{Mirror Match with Clustering (\emph{MMC})}
\label{alg-mmc}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$, $\alpha \in 
\mathbb{R}$
\State $M\gets \varnothing$
\State $F\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features
	\State $f_i\gets getFeatures(I_i)$
	\State $F\gets F \cup f_i$
\EndFor
\State $A\gets getAdjacencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets normalize(A)$
\ForAll{$(i,j) \in indices(A)$} \Comment Prune edges
    \If{$A_{norm}[i,j] < \alpha$}
        \State $A_{norm}[i,j] \gets 0$
    \EndIf
\EndFor
\State $P\gets cluster(A_{norm})$
\ForAll{$p \in P$} \Comment p is a set of feature points
    \State $M\gets M \cup getMatches(p, t, F)$ \Comment See 
    Algorithm~\ref{alg-getmatches}
\EndFor \\
\Return matches
\end{algorithmic}
%}
\end{algorithm}

To cluster the feature points, we construct a weighted graph with the 
feature points as nodes and their similarity score as normalized edge 
weights.  Given a set of $n$ features $F = {f_1, \ldots, f_n}$ and a 
similarity function $s$ we can define the matching function $M : (f_i, 
f_j) \rightarrow \mathbb{R}$ as follows:
\begin{equation*}
    M(a,b) = \begin{cases} s(f_i, f_j) & \mbox{if } f_i \neq f_j \\ 0 & 
    \mbox{otherwise}
	\end{cases}
\end{equation*}
The similarity function $s$ is based on the distance measure 
$d(f_i,f_j)$ (cosine similarity in the case of SIFT): $s(f_i, f_j) = 1 - 
d(f_i,f_j) / c$ where $c$ is a normalization constant normalizing the 
distance values between $0$ and $1$.
We obtain $A$, the normalized adjecency matrix of the fully connected 
similarity graph of feature points as follows:
\begin{equation*}
    A_{i,j} = \frac{M(f_i, f_j)}{\sum\limits_{f_i,f_j \in F} M(f_i, f_j) 
    / \left\vert F \right\vert ^ {2}}
\end{equation*}

In the literature there are various ways of clustering a graph according 
to different measures of what constitutes an optimal partitioning.  
Traditionally the most used clustering algorithms have been 
\emph{K-means} and spectral clustering, but in recent years a host of 
new algorithms have been proposed based on both Newman's concept of 
graph modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
well as information theoretical measures\footnote{See for example 
\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
in \cite{ronhovde2009}} just to mention a few.  On tests done using 
randomly generated graphs with a known partitioning, \cite{blondel2008}, 
\cite{rosvall2008} and \cite{ronhovde2009} perform markedly better than 
spectral clustering and \emph{K-means}\cite{lancichinetti2009}.

\begin{figure}[t]
    \centering
	\includegraphics[width=\textwidth]{images/MMC_graph}
    \caption{The partitioned feature graph. Each vertex represents a 
        feature point while lines indicate high similarity between 
        points.  A partition is a connected group with the same color.  
        The border color of each node indicates which image it belongs 
        to.  Zooming into a subsection of the graph (right part), the 
    various cluster sizes can be seen, ranging from hundreds of feature 
points to only two or three.}
    \label{fig:graph-view}
\end{figure}

We use the \emph{Louvain Method} \cite{blondel2008} for clustering 
feature points, since it is relatively fast and performs well 
\cite{lancichinetti2009}, does not require parameters 
\cite{blondel2008}, and does not emphasize partitions of equal size, as 
opposed to e.g.\ spectral clustering or k-means \cite{von2007}.

@@

We can optimize the modularity efficiently iteratively because the gain 
in modularity from moving a vertex $v_i$ to a cluster $C$ can be 
calculated easily:
\begin{equation*}
	\Delta Q = \left[ \frac{\sum_{in} + 2 k_{i,in}}{2m} - \
    \left(\frac{\sum_{tot} + k_i}{2m} \right)^2 \right]
	- \left[\frac{\sum_{in}}{2m} - \left(\frac{\sum_{tot}}{2m} \
	\right)^2 - \left( \frac{k_i}{2m} \right)^2 \right]
\end{equation*}
Here as in the definition of \emph{Modularity}, $m=\frac{1}{2} \sum_{ij} 
A_{ij}$, $k_i = \sum_j A_{ij}$ and $k_{i,in}$ is the summed weights of 
the edges from $v_i$ to vertices in $C$.  $\sum_{in}$ is the sum of the 
weights of the edges of $C$ while $\sum_{tot}$ is the sum of all edges 
with at least one vertex in $C$.  The algorithm works by assigning each 
vertex to its proper cluster and then for each iteration reassign 
vertices to neighboring clusters if it increases the \emph{Modularity}.  
These iterations are carried out until there are no reassignments in 
which case an initial clustering has been reached.  For graphs 
consisting of a large number of vertices, the algorithm can now be 
repeated, treating each of the resulting clusters as a vertex and 
repeating the same procedure.

@@

\begin{algorithm}[t]
\caption{Impl.\ of getMatches (\emph{from MMC algorithm})}
\label{alg-getmatches}
%\fontsize{10}{10}\selectfont
    \begin{algorithmic}
    \Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
    all features
    \State $edges \gets \left\{similarity(f_i, f_j) \mid getImg(f_i)
        \neq getImg(f_j) \wedge f_i, f_j \in p \right\}$
    \If{$\left\vert edges \right\vert > 1$}
        \State $matches \gets MMGet(p, t)$
    \ElsIf{$\left\vert edges \right\vert = 1$}
        \State $matches \gets RatioGet(p, features, t)$
    \Else
        \State $matches \gets \varnothing$
    \EndIf

    \Return matches
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
\caption{Ratio Match (\emph{Ratio})}
\label{alg-ratio}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches\gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NearestNeighbors(f_i, features_2)$
    \State $ratio \gets d(f_i, f_m) / d(f_i, f_n)$
	\If{$ratio < t$}
        \State $matches \gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
\Return $matches$
\end{algorithmic}
%}
\end{algorithm}

@@

\begin{algorithm}[h]
\caption{Isomatch (\emph{Isomatch})}
\label{alg-isomatch}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NNs(f_i, features_2)$
    \State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $partitions_1 \gets isodata(getPositions(features_1))$
\State $partitions_2 \gets isodata(getPositions(features_2))$
\State $P \gets getMatchMat(partitions_1, partitions_2, matches_{init})$
\ForAll{$(i,j) \in indices(P)$}
\If{$\left(P[i,j] > 5\right) \wedge \left(P[i,j] / sum(P_{i}) > 
0.5\right)$}
        \State $matches_{final} \gets getMatches(matches_{init}, 
        partitions_1, partitions_2, i, j)$
    \EndIf
\EndFor
\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

@@

The \emph{Isodata} algorithm is similar to \emph{K-means} with the 
addition that clusters are merged and split during the clustering 
process. In \emph{K-means} we randomly assign the data points to $k$ 
clusters to begin with and calculate their mean. Then for each data 
point we reassign it to the cluster closest in mean. We then calculate 
the mean of the updated clusters and reiterate the process until we have 
spent a set amount of iterations or no reassignments have taken place.  
\emph{Isodata} runs the same algorithm, but for each iteration it goes 
through the clusters and checks if any can be merged or split.  Clusters 
are split if their variance along an axis is beyond a threshold, while 
two clusters are merged if their distance is below another threshold.  
This approach is beneficial when we don't know from the outside the 
amount of clusters we expect to see. For example in the case of feature 
points it is not obvious before clustering them regionally how many 
regions we obtain for an image.

@@

\begin{algorithm}[h]
\caption{Spectral Match (\emph{Spectral})}
\label{alg-spectral}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final} \gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
	\State $f_m,f_n \gets get2NNs(f_i, features_2)$
	\State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $M \gets matrix(\left\vert matches_{init} \right\vert, \left\vert 
matches_{init} \right\vert)$
\ForAll{$m_i \in matches_{init}$}
	\ForAll{$m_j \in matches_{init}$}
		\If{$i = j$}
			\State $M \gets matchSimilarity(m_i)$
		\Else
			\State $M \gets affinity(m_i, m_j)$
		\EndIf
	\EndFor
\EndFor
\State $x^{*} \gets maximumEigenvector(M)$
\ForAll{$\left(m_i, x_i\right) \in \left(matches_{init}, x^{*}\right)$}
	\If{$x_i > t$}
		\State $matches_{final} \gets matchtes_{final} \cup m_i$
	\EndIf
\EndFor

\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

>>>>

In this chapter we will describe the algorithms used in the experiements
in detail. We start by introducing the proposed algorithms \emph{Mirror 
Match (MM)} and \emph{Mirror Match with Clustering (MMC)} and continue 
with the reference algorithms used later for comparison, \emph{Ratio}, 
\emph{Spectral} and \emph{Isomatch}.
%
\section{Proposed Matching Methods}
\label{S:MatchingMethods}
In this section we describe the two proposed algorithms for matching 
feature points \emph{Mirror Match
(MM)} and \emph{Mirror Match with Clustering (MMC)}. 
%

@@

%
\begin{figure}[h]
    \centering%
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_pitts_source}
            \caption{Source image pair}
            \label{fig:pitts_source}
        \end{subfigure}%
        ~ %
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_pitts_keypoints}
            \caption{Feature points}
            \label{fig:pitts_keypoints}
        \end{subfigure}%
        \\ %add desired spacing between images, e. g. ~, \quad, 
        %\qquad (or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match_off}
            \caption{\emph{Ratio}}
            \label{fig:unique}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, 
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match_with_pruned}
            \caption{\emph{MM} intermediate result}
            \label{fig:within}
        \end{subfigure}%
        \\ %add desired spacing between images, e. g. ~, \quad, 
        %\qquad (or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/mirror_match}
            \caption{\emph{MM} final result}
            \label{fig:without}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, 
        \begin{subfigure}[t]{0.5\columnwidth}
            \centering
            \includegraphics[width=0.95\columnwidth]{images/MMC_partition}
            \caption{\emph{MMC} Partition Example}
            \label{fig:pitts_partition}
        \end{subfigure}%
    \caption{Feature matching with \emph{MM} and \emph{MMC}. Dots 
        represent feature points; green/red lines indicate 
        correct/incorrect matches, respectively; black lines represent 
        edges in the feature graph.  (c) Result of \emph{Ratio} 
        matching.  (d) All matches found by \emph{MM}, including 
        intra-image matches.  (e) Final \emph{MM} result.  (f) Example 
    of a graph partition of feature points after clustering in 
\emph{MMC}, which includes similar feature points from different sides 
of a building.}%
    \label{fig:comparemirror}%
\end{figure}%
%
The central idea behind \emph{MM} is to match features of $n$ images by 
taking every feature from all $n$ images and matching them against every 
other feature from the same set. We then discard the correspondences 
that match two points within the same image. Algorithm~\ref{alg-mm} 
details the implementation of \emph{MM}.
%
\begin{algorithm}[htb]
\caption{Mirror Match (\emph{MM})}
\label{alg-mm}
{\fontsize{11}{11}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$, $matches_{final}\gets 
\varnothing$, $features\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Acquisition Stage
	\State $features\gets features \cup getFeatures(I_i)$
\EndFor
\ForAll{$f_i \in features$} \Comment Matching Stage
	\State $f_m,f_n \gets get2NNs(f_i, features \setminus 
	\left\{f_i\right\})$
    \State $ratio \gets d(f_i, f_m) / d(f_i, f_n)$
	\If{$ratio < t$}
		\State $matches_{init} \gets matches_{init} \cup \left(f_i, f_m\right)$
	\EndIf
\EndFor
\ForAll{$\left(f_i, f_j \right) \in matches_{init}$} \Comment Filter 
Stage
\If{$\left(f_j, f_i \right) \in matches_{init} \wedge getImg(f_i) \neq 
getImg(f_j) \wedge \left(f_j, f_i\right) \not\in matches_{final}$}
		\State $matches_{final} \gets (f_i, f_j)$
	\EndIf
\EndFor \\
\Return $matches_{final}$
\end{algorithmic}
}
\end{algorithm}
%
In the \emph{acquisition stage} we extract local feature points and 
obtain their descriptors from the set of images. The algorithm is not 
designed to make use of any feature extractor in particular, but for the
experiments (Chapter~\ref{C:Experiments}) the \emph{SIFT} feature 
extractor and descriptor has been used. In the \emph{matching} stage 
these features are matched using $k$-nearest neighbors.  The distance 
function $d(f_i, f_j)$ between features is calculated according to the 
feature descriptor. For \emph{SIFT} we use the cosine similarity between 
the two descriptor vectors, while other descriptors might use other 
measures\footnote{\emph{BRIEF} for example uses the hamming measure to 
speed up the computation.}.  For any given feature $f_i$ the two closest 
neighbors are returned, and we calculate the ratio between them as 
proposed in \cite{lowe2004sift}\footnote{In Part~\ref{ss:ratio} this 
calculation and other details of \emph{Ratio} will be discussed.}.  Any 
correspondence with a ratio above the threshold supplied will be 
discarded. Finally in the filter stage we check that matches are from 
different images and discard all matches that are not symmetric. That 
is, if a feature $f_i$, is best matched to another feature $f_j$, we 
would want for $f_j$ to be best matched to $f_i$ too.
This isn't always the case, for example if $f_i$ does not have a true 
correspondance in the other image, its closest neighbor $f_j$ might 
likely be better matched with a different feature point.

Figure~\ref{fig:comparemirror} illustrates the benefits of \emph{MM} 
using an example image pair from the \emph{Gallagher} dataset 
\cite{gallagher2008}.
With \emph{Ratio} (Figure~\ref{fig:unique}), many incorrect matches occur 
in the fence towards the bottom of the image.
When we match all feature points together, many of these incorrect 
matches are eliminated, because points in the fence match with other 
points in the fence in the same image (Figures~\ref{fig:within} and
\ref{fig:without}).
%

@@

%
In contrast to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering feature points by 
similarity. This process yields partitions of similar feature points 
that we can match using the same approach as \emph{MM}.  
Algorithm~\ref{alg-mmc} shows the implementation of \emph{MMC}.
%
\begin{algorithm}[htb]
\caption{Mirror Match with Clustering (\emph{MMC})}
\label{alg-mmc}
{\fontsize{11}{11}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$, $\alpha \in 
\mathbb{R}$
\State $M\gets \varnothing$, $F\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features
	\State $f_i\gets getFeatures(I_i)$
	\State $F\gets F \cup f_i$
\EndFor
\State $A\gets getAdjacencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets normalize(A)$
\ForAll{$(i,j) \in indices(A)$} \Comment Prune edges
    \If{$A_{norm}[i,j] < \alpha$}
        \State $A_{norm}[i,j] \gets 0$
    \EndIf
\EndFor
\State $P\gets cluster(A_{norm})$
\ForAll{$p \in P$} \Comment p is a set of feature points
    \State $M\gets M \cup getMatches(p, t, F)$ \Comment See 
    Algorithm~\ref{alg-getmatches}
\EndFor \\
\Return matches
\end{algorithmic}
}
\end{algorithm}
%
To cluster the feature points, we construct a weighted graph with the 
feature points as nodes and their similarity score as normalized edge 
weights.  Given a set of $n$ features $F = {f_1, \ldots, f_n}$ and a 
similarity function $s$ we can define the matching function $M : (f_i, 
f_j) \rightarrow \mathbb{R}$ as follows:
\begin{equation*}
    M(a,b) = \begin{cases} s(f_i, f_j) & \mbox{if } f_i \neq f_j \\ 0 & 
    \mbox{otherwise}
	\end{cases}
\end{equation*}
The similarity function $s$ is based on the distance measure 
$d(f_i,f_j)$ (cosine similarity in the case of SIFT): $s(f_i, f_j) = 1 - 
d(f_i,f_j) / c$ where $c$ is a normalization constant normalizing the 
distance values between $0$ and $1$.
We obtain $A$, the normalized adjecency matrix of the fully connected 
similarity graph of feature points as follows:
\begin{equation*}
    A_{i,j} = \frac{M(f_i, f_j)}{\sum\limits_{f_i,f_j \in F} M(f_i, f_j) 
    / \left\vert F \right\vert ^ {2}}
\end{equation*}
%
Traditionally the most used clustering algorithms have been 
\emph{K-means} and spectral clustering, but in recent years a host of 
new algorithms have been proposed based on both Newman's concept of 
\emph{Modularity}\footnote{Introduced in \cite{girvan2002}, discussed in 
\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
well as information theoretical measures\footnote{See for example 
\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
in \cite{ronhovde2009}} just to mention a few.  On tests done using 
randomly generated graphs with a known partitioning, \cite{blondel2008}, 
\cite{rosvall2008} and \cite{ronhovde2009} perform markedly better than 
spectral clustering and \emph{K-means}\cite{lancichinetti2009}.
%
\begin{figure}[t]
    \centering
	\includegraphics[width=\textwidth]{images/MMC_graph}
    \caption{The partitioned feature graph. Each vertex represents a 
        feature point while lines indicate high similarity between 
        points.  A partition is a connected group with the same color.  
        The border color of each node indicates which image it belongs 
        to.  Zooming into a subsection of the graph (right part), the 
    various cluster sizes can be seen, ranging from hundreds of feature 
points to only two or three.}
    \label{fig:graph-view}
\end{figure}
%
We use the \emph{Louvain Method} \cite{blondel2008} for clustering 
feature points, since it is relatively fast and performs well 
\cite{lancichinetti2009}, does not require parameters 
\cite{blondel2008}, and does not emphasize partitions of equal size, as 
opposed to e.g.\ spectral clustering or k-means \cite{von2007}.

@@

We can optimize the \emph{Modularity} efficiently iteratively because 
the gain in \emph{Modularity} from moving a vertex $v_i$ to a cluster 
$C$ can be calculated easily:
\begin{equation*}
	\Delta Q = \left[ \frac{\sum_{in} + 2 k_{i,in}}{2m} - \
    \left(\frac{\sum_{tot} + k_i}{2m} \right)^2 \right]
	- \left[\frac{\sum_{in}}{2m} - \left(\frac{\sum_{tot}}{2m} \
	\right)^2 - \left( \frac{k_i}{2m} \right)^2 \right]
\end{equation*}
Here as in the definition of \emph{Modularity}, $m=\frac{1}{2} \sum_{ij} 
A_{ij}$, $k_i = \sum_j A_{ij}$ and $k_{i,in}$ is the summed weights of 
the edges from $v_i$ to vertices in $C$.  $\sum_{in}$ is the sum of the 
weights of the edges of $C$ while $\sum_{tot}$ is the sum of all edges 
with at least one vertex in $C$.  The algorithm works by assigning each 
vertex to its proper cluster and then for each iteration reassign 
vertices to neighboring clusters if it increases the \emph{Modularity}.  
These iterations are carried out until there are no reassignments in 
which case an initial clustering has been reached.  For graphs 
consisting of a large number of vertices, the algorithm can now be 
repeated, treating each of the resulting clusters as a vertex and 
repeating the same procedure.

@@

\begin{algorithm}[t]
\caption{Impl.\ of getMatches (\emph{from MMC algorithm})}
\label{alg-getmatches}
{\fontsize{11}{11}\selectfont
    \begin{algorithmic}
    \Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
    all features
    \State $edges \gets \left\{similarity(f_i, f_j) \mid getImg(f_i)
        \neq getImg(f_j) \wedge f_i, f_j \in p \right\}$
    \If{$\left\vert edges \right\vert > 1$}
        \State $matches \gets MMGet(p, t)$
    \ElsIf{$\left\vert edges \right\vert = 1$}
        \State $matches \gets RatioGet(p, features, t)$
    \Else
        \State $matches \gets \varnothing$
    \EndIf

    \Return matches
    \end{algorithmic}
}
\end{algorithm}

@@

\begin{algorithm}[t]
\caption{Ratio Match (\emph{Ratio})}
\label{alg-ratio}
{\fontsize{11}{11}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches\gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NearestNeighbors(f_i, features_2)$
    \State $ratio \gets d(f_i, f_m) / d(f_i, f_n)$
	\If{$ratio < t$}
        \State $matches \gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
\Return $matches$
\end{algorithmic}
}
\end{algorithm}

@@

\begin{algorithm}[h]
\caption{Isomatch (\emph{Isomatch})}
\label{alg-isomatch}
{\fontsize{11}{11}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$, $matches_{final}\gets 
\varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NNs(f_i, features_2)$
    \State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $partitions_1 \gets isodata(getPositions(features_1))$
\State $partitions_2 \gets isodata(getPositions(features_2))$
\State $P \gets getMatchMat(partitions_1, partitions_2, matches_{init})$
\ForAll{$(i,j) \in indices(P)$}
\If{$\left(P[i,j] > 5\right) \wedge \left(P[i,j] / sum(P_{i}) > 
0.5\right)$}
        \State $matches_{final} \gets getMatches(matches_{init}, 
        partitions_1, partitions_2, i, j)$
    \EndIf
\EndFor
\Return $matches_{final}$
\end{algorithmic}
}
\end{algorithm}

@@

The \emph{Isodata} algorithm is similar to \emph{K-means} with the 
addition that clusters are merged and split during the clustering 
process. In \emph{K-means} we randomly assign the data points to $k$ 
clusters to begin with and calculate their mean. Then for each data 
point we reassign it to the cluster closest in mean. We then calculate 
the mean of the updated clusters and reiterate the process until we have 
spent a set amount of iterations or no reassignments have taken place.  
\emph{Isodata} runs the same algorithm, but for each iteration it goes 
through the clusters and checks if any can be merged or split.  Clusters 
are split if their variance along an axis is beyond a threshold, while 
two clusters are merged if their distance is below another threshold.  
This approach is beneficial when we do not know from the outside the 
amount of clusters we expect to see. For example in the case of feature 
points it is not obvious before clustering them regionally how many 
regions we obtain for an image.

@@

\begin{algorithm}[h]
\caption{Spectral Match (\emph{Spectral})}
\label{alg-spectral}
{\fontsize{11}{11}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$, $matches_{final} \gets 
\varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
	\State $f_m,f_n \gets get2NNs(f_i, features_2)$
	\State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $M \gets matrix(\left\vert matches_{init} \right\vert, \left\vert 
matches_{init} \right\vert)$
\ForAll{$m_i \in matches_{init}$}
	\ForAll{$m_j \in matches_{init}$}
		\If{$i = j$}
			\State $M \gets matchSimilarity(m_i)$
		\Else
			\State $M \gets affinity(m_i, m_j)$
		\EndIf
	\EndFor
\EndFor
\State $x^{*} \gets maximumEigenvector(M)$
\ForAll{$\left(m_i, x_i\right) \in \left(matches_{init}, x^{*}\right)$}
	\If{$x_i > t$}
		\State $matches_{final} \gets matchtes_{final} \cup m_i$
	\EndIf
\EndFor

\Return $matches_{final}$
\end{algorithmic}
}
\end{algorithm}

\emph{Spectral} (Algorithm~\ref{alg-spectral}) is based on the spectral 
matching algorithm proposed by Leordeanu and Herbert 
\cite{leordeanu2005spectral}.  The central idea behind spectral matching 
is that the set of actual correspondences of feature points between two 
images should be pairwise consistent. That is, given two feature points 
that are geometrically close in one image, for most cases we would 
assume that the closest correspondences of these two points in another 
image would be similarly close.  Based on this assumption we want to 
pick out a set of matches $T \subset S$ where $S$ is the initial set of 
possible matches such that the matches in $T$ are as pairwise consistent 
as possible. If we let $x$ be a binary vector corresponding in length to 
the set of initial matches $S$, where $x_i = 1$ if $m_i \subset T$, then 
we can formulate a subset-score as:
\begin{equation*}
	S = \sum_{m_i, m_j \in T} M(m_i, m_j) = x^TMx
\end{equation*}
The solution $x^{*}$ that maximizes this equation is then $x^{*} = 
argmax(x^TMx)$. By relaxing the constraints on $x$ such that it can take
values in the range $\left[0, 1\right]$ we can interpret $x_i$ as the
association of $m_i$ with the optimal subset of matches $T$. Because we 
only worry about the relative values of $x$ we can fix the norm of $x$ 
to $1$, in which case by Raleigh's ratio theorem, the $x^{*}$ that 
optimizes $x^TMx$ is the principal eigenvector of $M$.
